Where Do Events Come From?

### 2018-07-18 paper.tex
...

\begin{abstract}
In reactive and event-based systems, execution is guided by an external
environment which produces inputs and consumes output events to and from the
applications.
Reactive languages provide dedicated syntax and semantics to deal with events
and greatly simplify the programming experience in this domain.
%
Nevertheless, the environment is typically prefabricated in a host language and
the very central concept of events is implemented externally to the reactive
language.
This separation exposes the environment as a rigid system component that
evolves in separate from the application.
%
In this work, we propose an interrupt handler primitive for a reactive language
targeting embedded systems in order to take control of the whole event loop,
from the hardware input source and back to the hardware output sink.
%
We propose the new primitive in the context of the structured reactive language
\CEU and discuss how it synergizes with the structured paradigm to provide
automatic standby for applications and hot swapping for drivers during runtime.
\end{abstract}

@@

Reactive applications interact continuously and in real time with the external
world through input and output devices (e.g., buttons, displays, timers, etc.).
%
These interactions are typically represented as input events flowing from the
devices to the application and as output events flowing from the application to
the devices.
%
As illustrated in Figure~\ref{fig.loop}, devices can be encapsulated as a
single component, \emph{the environment}, which controls the system execution
in an event loop:
the application sits idle waiting for an input;
the environment awakes the application on the occurrence of an input;
the application reacts to the input and generates back one or more outputs;
the application becomes idle and the loop restarts.

@@

The environment is typically implemented in a host language (e.g., C) and
controls the main event loop, invoking entry points in the reactive language
runtime on the occurrence of inputs and also receiving output calls, both
through a documented API.
%
As examples, Esterel~\cite{esterel.ieee91} relies on C for passing events
between the environment and the running program~\cite{esterel.book.compiling},
while Elm~\cite{frp.elm} uses the concept of \emph{ports}, which allows sending
out values to JavaScript as commands and listening for values as
subscriptions~\cite{frp.elm.ports}.

The event-based separation between the application and the environment is
arguably inevitable and happens at the right level of abstraction, since it
connects the application with the operating system resources through a
non-invasive API.
%
However, it involves two languages and the programmer may have to unnecessarily
deal with multiple syntaxes, incompatible type systems, and different address
spaces.
%
Furthermore, in the context of embedded systems, a proper host OS may even be
absent or lacking enough drivers, which requires more low-level intervention
from the application.

@@

>>>>

\begin{abstract}
In reactive and event-based systems, execution is guided by an external
environment which generates inputs to the application and consumes outputs from
it.
Reactive languages provide dedicated syntax and semantics to deal with events
and greatly simplify the programming experience in this domain.
%
Nevertheless, the environment is typically prefabricated in a host language and
the very central concept of events is implemented externally to the reactive
language.
%
In this work, we propose an interrupt handler primitive for a reactive language
targeting embedded systems in order to take control of the whole event loop,
from a sensor input source and back to an actuator output.
%
We propose the new primitive in the context of the structured reactive language
\CEU and discuss they synergize to provide automatic standby for applications
and hot swapping of device drivers, while avoiding race conditions.
\end{abstract}

@@

Reactive applications interact continuously and in real time with the external
world through sensors and actuators (e.g., buttons, displays, timers, etc.).
%
These interactions are typically represented as input events flowing from the
devices to the application and as output events flowing from the application to
the devices.
%
As illustrated in Figure~\ref{fig.loop}, devices can be encapsulated as a
single component, \emph{the environment}, which controls the system execution
in an event loop:
the application sits idle waiting for an input;
the environment awakes the application on the occurrence of an input;
the application reacts to the input and generates back one or more outputs;
the application becomes idle and the loop restarts.

@@

The environment is typically implemented in a host language (e.g., C) and
controls the main event loop, invoking entry points in the reactive language
runtime on the occurrence of inputs and also receiving output calls, both
through a documented API.
%
As examples, Esterel~\cite{esterel.ieee91} relies on C for passing events
between the environment and the running program~\cite{esterel.book.compiling},
while Elm~\cite{frp.elm} uses the concept of \emph{ports}, which allows sending
out values to JavaScript in commands and listening for values in
subscriptions~\cite{frp.elm.ports}.

%This separation exposes the environment as a rigid system component that
%evolves in separate from the application.
The event-based separation between the application and the environment is
arguably inevitable and also happens at the right level of abstraction, since
it connects the application with the operating system resources through a
non-invasive API.
%
However, it involves two languages and the programmer may have to deal with
multiple syntaxes, incompatible type systems, and different address spaces.
%
Furthermore, in the context of embedded systems, a proper host OS may even be
absent or lacking enough device drivers, which requires more low-level
intervention from the application.

@@

We present two uses cases that take advantage of the tight integration between
the language and environment:
